### 2022-07-14 draft.tex
Deploying to gh-pages from @ FrancisBanville/ms_maxent_networks@3d6d340641bce3d032a9391b700ea89c923f4a03 🚀

Statistical and mathematical models can help fill many gaps in our
knowledge about species interactions. Two complementary types of models
have been developed in network ecology for this purpose. On one hand,
predictive models can partially alleviate the Eltonian shortfall, which
describes our current lack of knowledge on food webs and other
ecological networks (Hortal et al. 2015). A variety of such models have
recently been developed using machine learning and other statistical
tools, most of which are presented in Strydom et al. (2021). On the
other hand, null models help us identify potential ecological mechanisms
that drive species interactions. They do so by comparing empirical data
with an unbiased distribution of measures generated using a set of rules
that exclude the mechanism of interest (Fortuna and Bascompte 2006;
Delmas et al. 2019). Both types of models are frequently topological,
i.e.~they often predict the adjacency matrix or specific measures of
network structure without taking into account species' identity.
According to Strydom et al. (2021), these topological models could be
used to make better predictions of pairwise species interactions by
constraining the space of feasible networks.

@@

Despite its extensive use in graph and network theories (e.g., Park and
Newman 2004; van der Hoorn, Lippner, and Krioukov 2018), MaxEnt has in
comparison been little used in network ecology. The very nature of
ecological networks (directed simple graphs frequently having
self-loops) makes the mathematical optimization of maximum entropy graph
models more complicated than with many other types of (non-ecological)
networks. MaxEnt has nevertheless been used to predict the degree
distribution of bipartite ecological networks from the number of species
and the number of interactions (Williams 2011) and to predict
interaction strengths between species pairs using their relative
abundances within an optimal transportation theory regularized with
information entropy (Stock, Poisot, and De Baets 2021). However, to the
best of our knowledge, MaxEnt has never been used to predict food-web
structure directly, even though food webs are among the most documented
and widespread ecological networks.

In this contribution, we used two complementary approaches to predict
the structure of food webs using the principle of maximum entropy. We
then compared our predictions against empirical data and null and
neutral models commonly used in network ecology. The first approach
consists in deriving constrained probability distributions of given
network properties directly. We derived the joint degree distribution (a
probability distribution) of maximum entropy using only the number of
species \(S\) and the number of interactions \(L\) as state variables.
Then, we predicted the degree distribution of maximum entropy directly
from the joint degree distribution since the first is the sum of the
marginal distributions of the second (a species' degree is the sum of
its in and out-degrees). Because of the scarcity of empirical data on
the number of links in ecological networks, in Box 2 we present a method
to predict \(L\) from \(S\), thus allowing the prediction of the joint
degree distribution from \(S\) solely. In turn, the second approach
consists in finding, using different constraints, the adjacency matrix
of maximum entropy from which network properties can be measured. To do
so, we used a flexible and heuristic approach based on simulated
annealing to find networks \emph{close} to maximum entropy. As discussed
above, our choice of algorithm stands from the very nature of food webs
(i.e., simple directed networks allowing self-loops) that makes the
analytical derivation of a maximum entropy graph model difficult. We
first built our type I MaxEnt network model constrained by the
connectance of the network (i.e., the ratio \(L/S^2\)). A comparison of
this model against empirical data indicated that connectance alone was
not sufficient to predict many aspects of network structure. For this
reason, we built our type II MaxEnt network model, which instead uses
the whole joint degree sequence as a constraint. Overall, we found that
this second model was much better at predicting food-web structure than
the one based on connectance.

@@

The principle of maximum entropy is a mathematical method of finding
probability distributions, strongly rooted in statistical mechanics and
information theory (Jaynes 1957a, 1957b; Harremoës and Topsøe 2001).
Starting from a set of constraints given by prior knowledge of a system
(i.e., what we call state variables), this method helps us find
least-biased probability distributions subject to the constraints. These
probability distributions are guaranteed to be unique given our prior
knowledge and represent the most we can say about a system without
making more assumptions. For example, if the only thing we know about a
biological community is its average number of individuals per species,
the least-biased inference we could make on its species abundance
distribution is the exponential distribution (Frank and Smith 2011;
Harte and Newman 2014). However, this does not imply that this
distribution will be the best fit to empirical data. The challenge is to
find the right set of constraints that would best reproduce
distributions found in nature.

@@

where \(g_i\) is the mathematical formulation of the constraint \(i\)
and \(c_i\), its value. Note that \(F\) is just Shannon's entropy to
which we added terms that each sums to zero (\(g_i = c_i\)). \(F\) is
maximized by setting to \(0\) its partial derivative with respect to
\(p(n)\). We will show how this can be done when we derive the joint
degree distribution analytically from the number of species and the
number of links in food webs.

@@

where \(s_i\) are the relative singular values
(\(s_i = \sigma_i / \sum_{i = 1}^R \sigma_i\), where \(\sigma_i\) are
the singular values). Following Strydom, Dalla Riva, and Poisot (2021),
we standardized this measure with the rank \(R\) of the matrix (i.e.,
\(J / \ln(R)\)) to account for the difference in dimensions between
networks (Pielou's evenness; Pielou 1975). In a following section, we
will show how SVD entropy can be used to predict a network of maximum
entropy (i.e., of maximum complexity) heuristically.

@@

We predicted both types of null networks for all empirical networks in
our complete dataset (\(n = 257\)). We converted all probabilistic
networks to Boolean networks by generating \(100\) random Boolean
networks for each of these probabilistic webs. Then, we counted the
number of times each interaction was sampled, and kept the \(L\) entries
that were drawn the most amount of time, with \(L\) given by the number
of links in each food web. This ensured that the resulting null networks
had the same number of interactions as their empirical counterparts.

@@

We predicted the joint degree distribution of maximum entropy for each
food web in our complete dataset, i.e.~using their numbers of species
and numbers of links as state variables. We then sampled one realization
of the degree sequence for each network using the probabilities given by
the joint degree distribution. In fig.~\ref{fig:joint_dd} (left panels),
we show the relationship between \(k_{out}\) and \(k_{in}\) standardized
by the number of species in their networks, for empirical and maximum
entropy joint degree distributions. We see that our model predicts a
greater number of generalist species than empirical food webs (i.e.,
relative values of \(k_{out}\) and \(k_{in}\) closer to \(1\)). However,
plotting the difference between predicted and empirical values for each
species gives a different perspective. The right panel of
fig.~\ref{fig:joint_dd} presents these differences when species are
ordered by their total degree in their network (i.e., by the sum of
their in and out-degrees). Indeed, our predicted joint degree sequences
have the same number of species as their empirical counterparts, but
they are species agnostic; in other words, instead of predicting a value
for each species directly, we predicted the entire joint degree sequence
without taking into account species' identity. When we associate
predictions and empirical data according to their rank in total degrees,
we see that species predicted to be have a higher generality (number of
preys) generally have a lower vulnerability (number of preys) than what
is observed (and conversely). In fig.~S1, we show how these differences
change when species are instead ordered by their out-degrees (left
panel) and in-degrees (right panel), respectively.

@@

The degree distribution \(p(k)\) represents the probability that a
species has \(k\) links in a food web, with \(k = k_{in} + k_{out}\). It
can thus be directly obtained from the joint degree distribution:

@@

In fig.~S2, we show that the degree distribution of maximum entropy,
given \(S\) and \(L\), predicts very low probabilities that a species
will be isolated in its food web (\emph{i.e.}, having \(k = 0\)). As
MacDonald, Banville, and Poisot (2020) pointed out, the size of food
webs should at least be of \(S-1\) links, since a lower number would
yield isolated species, i.e.~species without any predators or preys. Our
results show that, under our purely information-theoretic model, the
probability that a species is isolated is quite high below this
threshold. The expected proportion of isolated species rapidly declines
by orders of magnitude with increasing numbers of species and links.

@@

\hypertarget{box-2---working-with-predicted-numbers-of-links}{%
\section{Box 2 - Working with predicted numbers of
links}\label{box-2---working-with-predicted-numbers-of-links}}

Our models need information on the number of species and the number of
links. However, since the later is rarely estimated empirically,
ecologists might need to use predictive methods to estimate the total
number of links in a food web.

We used the flexible links model of MacDonald, Banville, and Poisot
(2020) to predict the number of interactions from the number of species.
The flexible links model, in contrast to other predictive models of the
number of links, incorporates meaningful ecological constraints into the
prediction of \(L\), namely the minimum \(S-1\) and maximum \(S^2\)
numbers of interactions in food webs. It estimates the proportion of the
\(S^2 - (S - 1)\) \emph{flexible links} that are realized. More
precisely, this model states that the number of \emph{realized} flexible
links \(L_{FL}\) in a food web represents the number of realized
interactions above the minimum (i.e., \(L = L_{FL} + S - 1\)) and is
obtained from a beta-binomial distribution with \(S^2 - (S - 1)\) trials
and parameters \(\alpha = \mu e^\phi\) and \(\beta = (1 - \mu) e^\phi\):

@@

\begin{figure}
\hypertarget{fig:degree_dist_fl}{%
\centering
\includegraphics{figures/maxent_degree_dist_fl.png}
\caption{(a) Probability density of the predicted mean degree for
different values of species richness. The number of links was predicted
using the flexible links model fitted to all empirical networks in our
complete dataset. (b) Degree distributions of maximum entropy for a
network of \(27\) species and different numbers of links. The numbers of
links correspond to the lower and upper bounds of the \(67%
\), \(89%
\), and \(97%
\) percentile intervals (PI), as well as the median, of the
counterfactuals of the flexible links model.}\label{fig:degree_dist_fl}
}
\end{figure}

@@

We define networks of maximum entropy as the configuration of the
adjacency matrix with the highest SVD entropy under a set of
constraints. As mentioned in Box 1, we used the SVD entropy as our
measure of entropy since it has been shown to be a reliable measure of
food-web complexity (Strydom, Dalla Riva, and Poisot 2021), in addition
to having the required properties of a proper measure of information
entropy. We thus seek to find the network with the highest complexity,
or randomness, that exactly reproduces specified constraints on its
structure. Our method is in contrast with maximum entropy graph models
that predict a probability distribution on networks under soft or hard
constraints Cimini et al. (2019). We believe our approach to be more
flexible, easier to compute, while allowing direct comparisons of
empirical food webs with more complex networks with similar structure.

@@

We compared different measures of the structure of empirical food webs
with the ones of null, neutral, and MaxEnt networks
tbl.~\ref{tbl:measures_all}, and \ref{tbl:measures_abund}. For instance,
we measured the nestedness \(\rho\) according to Staniczenko, Kopp, and
Allesina (2013), i.e.~using the spectral radius of the adjacency matrix.
Nestedness indicates how much the diet of specialist species is a subset
of the one of generalists, and is strongly associated with network
modularity (Fortuna et al. 2010). We also measured the maximum trophic
level \emph{maxtl}, network diameter \emph{diam} (i.e., the longest of
the shortest paths between all species pairs; Albert and Barabasi 2002),
the average maximum similarity between species pairs \emph{MxSim}, the
proportion of cannibal species \emph{Cannib} (i.e., the proportion of
self loops), and the proportion of omnivorous species \emph{Omniv}
(i.e., species whose preys are of different trophic levels).
\emph{MxSim}, \emph{Cannib}, and \emph{Omniv} are more deeply defined in
Williams and Martinez (2000).

@@

Moreover, we found that empirical networks are close to their maximum
entropy considering a fixed joint degree sequence (fig.~S3). Empirical
networks had an SVD entropy of \(0.89 \pm 0.04\), compared to type II
MaxEnt networks which had values of \(0.94 \pm 0.03\). The relationship
between the SVD entropy of empirical food webs and their maximum entropy
configuration is plotted in the last panel of fig.~\ref{fig:measures}.
We also found no relationship between the difference in SVD entropy
between empirical and type II MaxEnt networks and species richness, the
number of links, or connectance (fig.~S4). Similarly, we found no
correlation between the difference in SVD entropy and the Jaccard
distance of empirical and type II MaxEnt networks.

@@

The principle of maximum entropy is a robust mathematical method of
finding least-biased probability distributions that have some desired
properties given by prior knowledge on a system. We first used MaxEnt to
predict the joint degree distribution and the degree distribution of
maximum entropy given known numbers of species and of interactions. We
found that the resulting joint degree distributions were more symmetric
than the ones of empirical food webs, which suggests that other
constraints might be needed to improve those predictions. We also used
MaxEnt to predict networks of maximum entropy with a specified
structure. These networks are the most complex, or random, given the
constraints used. Likewise, we found that knowledge of species richness
and of the number of links were not sufficient to reproduce many aspects
of network structure found in nature. However, a model based on the
entire joint degree sequence, i.e.~on the number of preys and predators
for each species, gave more convincing results. Our type II MaxEnt
network model yielded better or similar predictions than the type II
null model, also based on the joint degree sequence, for most measures
considered including the motifs profile. This suggests that the joint
degree sequence drives many aspects of network structure. Indeed,
considering our findings that empirical networks are close to maximum
entropy for a given joint degree sequence, our results suggest that
food-web topology is configured almost entirely randomly around these
marginal numbers of predators and preys.

Our method and results could be used for different purposes. First, they
could be used as first-order approximations of network structure when
only state variables are known. This could prove useful when predicting
network structure at large spatial scales, where few ecological
information is known at that scale. Second, they could be used as
informative priors in Bayesian analyses of the structure of ecological
networks. Third, they could be used to make better predictions of
pairwise species interactions by constraining the space of feasible
networks, as discussed in Strydom et al. (2021). Finally, they could be
used as alternative null models of ecological networks to better
understand ecological mechanisms driving food-web structure. In that
case, our model might need to be slightly adapted to give a probability
distribution of Boolean networks (in contrast with point estimates of
maximum entropy networks).

@@

>>>>

Statistical and mathematical models can help fill many gaps in our
knowledge about species interactions. Predictive and null models are two
complementary types of models that have been developed in network
ecology for this purpose. On one hand, predictive models can partially
alleviate the Eltonian shortfall, which describes our current lack of
knowledge on food webs and other ecological networks (Hortal et al.
2015). A variety of such models have recently been developed using
machine learning and other statistical tools, most of which are
presented in Strydom et al. (2021). On the other hand, null models help
us identify potential ecological mechanisms that drive species
interactions. They do so by comparing empirical data with an unbiased
distribution of measures generated using a set of rules that exclude the
mechanism of interest (Fortuna and Bascompte 2006; Delmas et al. 2019).
Both types of models are frequently topological, i.e.~they often predict
the adjacency matrix or specific measures of network structure without
taking into account species' identity. According to Strydom et al.
(2021), these topological models could be used to make better
predictions of pairwise species interactions by constraining the space
of feasible networks.

@@

Despite its extensive use in graph and network theories (e.g., Park and
Newman 2004; van der Hoorn, Lippner, and Krioukov 2018), MaxEnt has in
comparison been little used in network ecology. Like other real system
networks, ecological networks are represented mathematically as graphs.
However, the very nature of ecological networks (directed simple graphs
frequently having self-loops) makes the mathematical optimization of
maximum entropy graph models more complicated than with many other types
of (non-ecological) networks. MaxEnt has nevertheless been used to
predict the degree distribution of bipartite ecological networks from
the number of species and the number of interactions (Williams 2011) and
to predict interaction strengths between species pairs using their
relative abundances within an optimal transportation theory regularized
with information entropy (Stock, Poisot, and De Baets 2021). However, to
the best of our knowledge, MaxEnt has never been used to predict
food-web structure directly, even though food webs are among the most
documented and widespread ecological networks.

In this contribution, we used two complementary approaches to predict
the structure of food webs using the principle of maximum entropy. We
then compared our predictions against empirical data and null and
neutral models commonly used in network ecology. The first approach
consists in deriving constrained probability distributions of given
network properties directly. We derived the joint degree distribution (a
probability distribution) of maximum entropy using only the number of
species \(S\) and the number of interactions \(L\) as state variables.
Then, we predicted the degree distribution of maximum entropy directly
from the joint degree distribution since the first is the sum of the
marginal distributions of the second (a species' degree is the sum of
its in and out-degrees). Because of the scarcity of empirical data on
the number of interactions in ecological networks, we present a method
to predict \(L\) from \(S\) (Box 2), thus allowing the prediction of the
joint degree distribution from \(S\) solely. In turn, the second
approach consists in finding, using different constraints, the adjacency
matrix of maximum entropy from which network properties can be measured.
To do so, we used a flexible and heuristic approach based on simulated
annealing to find networks \emph{close} to maximum entropy. As discussed
above, our choice of algorithm stands from the very nature of food webs
(i.e.~simple directed networks allowing self-loops) that makes the
analytical derivation of a maximum entropy graph model difficult. We
first built our type I MaxEnt network model constrained by the
connectance of the network (i.e.~the ratio \(L/S^2\)). A comparison of
this model against empirical data indicated that connectance alone was
not sufficient to predict many aspects of network structure. For this
reason, we built our type II MaxEnt network model, which instead uses
the whole joint degree sequence as a constraint. Overall, we found that
this second model was much better at predicting food-web structure than
the one based on connectance.

@@

The principle of maximum entropy is a mathematical method of finding
probability distributions, strongly rooted in statistical mechanics and
information theory (Jaynes 1957a, 1957b; Harremoës and Topsøe 2001).
Starting from a set of constraints given by prior knowledge of a system
(i.e.~what we call state variables), this method helps us find
least-biased probability distributions subject to the constraints. These
probability distributions are guaranteed to be unique given our prior
knowledge and represent the most we can say about a system without
making more assumptions. For example, if the only thing we know about a
biological community is its average number of individuals per species,
the least-biased inference we could make on its species abundance
distribution is the exponential distribution (Frank and Smith 2011;
Harte and Newman 2014). However, this does not imply that this
distribution will be the best fit to empirical data. The challenge is to
find the right set of constraints that would best reproduce
distributions found in nature.

@@

where \(g_i\) is the mathematical formulation of the constraint \(i\)
and \(c_i\), its value. Note that \(F\) is just Shannon's entropy to
which we added terms that each sums to zero (\(g_i = c_i\)). \(F\) is
maximized by setting to \(0\) its partial derivative with respect to
\(p(n)\). We will show how this can be done when we derive the joint
degree distribution analytically from the number of species and the
number of interactions in food webs.

@@

where \(s_i\) are the relative singular values
(\(s_i = \sigma_i / \sum_{i = 1}^R \sigma_i\), where \(\sigma_i\) are
the singular values). Following Strydom, Dalla Riva, and Poisot (2021),
we standardized this measure with the rank \(R\) of the matrix
(i.e.~\(J / \ln(R)\)) to account for the difference in dimensions
between networks (\emph{sensu} Pielou's evenness; Pielou 1975). In a
following section, we will show how SVD entropy can be used to predict a
network of maximum entropy (i.e.~of maximum complexity) heuristically.

@@

We predicted both types of null networks for all empirical networks in
our complete dataset (\(n = 257\)). We converted all probabilistic
networks to Boolean networks by generating \(100\) random Boolean
networks for each of these probabilistic webs. Then, we counted the
number of times each interaction was sampled, and kept the \(L\) entries
that were drawn the most amount of time, with \(L\) given by the number
of interactions in each food web. This ensured that the resulting null
networks had the same number of interactions as their empirical
counterparts.

@@

We predicted the joint degree distribution of maximum entropy for each
food web in our complete dataset, i.e.~using their numbers of species
and numbers of interactions as state variables. We then sampled one
realization of the degree sequence for each network using the
probabilities given by the joint degree distribution. In
fig.~\ref{fig:joint_dd} (left panels), we show the relationship between
\(k_{out}\) and \(k_{in}\) standardized by the number of species in
their networks, for empirical and maximum entropy joint degree
distributions. We see that our model predicts higher values of
generality and vulnerability compared to empirical food webs
(i.e.~relative values of \(k_{out}\) and \(k_{in}\) closer to \(1\)).
However, plotting the difference between predicted and empirical values
for each species gives a different perspective. The right panel of
fig.~\ref{fig:joint_dd} presents these differences when species are
ordered by their total degree in their network (i.e.~by the sum of their
in and out-degrees). Indeed, our predicted joint degree sequences have
the same number of species as their empirical counterparts, but they are
species agnostic; in other words, instead of predicting a value for each
species directly, we predicted the entire joint degree sequence without
taking into account species' identity. When we associate predictions and
empirical data according to their rank in total degrees, we see that
species predicted to be have a higher generality (number of preys)
generally have a lower vulnerability (number of preys) than what is
observed (and conversely). In fig.~S1, we show how these differences
change when species are instead ordered by their out-degrees (left
panel) and in-degrees (right panel), respectively.

@@

The degree distribution \(p(k)\) represents the probability that a
species has \(k\) interactions in a food web, with
\(k = k_{in} + k_{out}\). It can thus be directly obtained from the
joint degree distribution:

@@

In fig.~S2, we show that the degree distribution of maximum entropy,
given \(S\) and \(L\), predicts very low probabilities that a species
will be isolated in its food web (i.e.~having \(k = 0\)). As MacDonald,
Banville, and Poisot (2020) pointed out, the size of food webs should at
least be of \(S-1\) interactions, since a lower number would yield
isolated species, i.e.~species without any predators or preys. Our
results show that, under our purely information-theoretic model, the
probability that a species is isolated is quite high below this
threshold. The expected proportion of isolated species rapidly declines
by orders of magnitude with increasing numbers of species and
interactions.

@@

\hypertarget{box-2---working-with-predicted-numbers-of-interactions}{%
\section{Box 2 - Working with predicted numbers of
interactions}\label{box-2---working-with-predicted-numbers-of-interactions}}

Our models need information on the number of species and the number of
interactions. However, since the later is rarely estimated empirically,
ecologists might need to use predictive methods to estimate the total
number of interactions in a food web.

We used the flexible links model of MacDonald, Banville, and Poisot
(2020) to predict the number of interactions from the number of species.
The flexible links model, in contrast to other predictive models of the
number of interactions, incorporates meaningful ecological constraints
into the prediction of \(L\), namely the minimum \(S-1\) and maximum
\(S^2\) numbers of interactions in food webs. It estimates the
proportion of the \(S^2 - (S - 1)\) \emph{flexible links} that are
realized. More precisely, this model states that the number of
\emph{realized} flexible links (or interactions) \(L_{FL}\) in a food
web represents the number of realized interactions above the minimum
(i.e.~\(L = L_{FL} + S - 1\)) and is obtained from a beta-binomial
distribution with \(S^2 - (S - 1)\) trials and parameters
\(\alpha = \mu e^\phi\) and \(\beta = (1 - \mu) e^\phi\):

@@

\begin{figure}
\hypertarget{fig:degree_dist_fl}{%
\centering
\includegraphics{figures/maxent_degree_dist_fl.png}
\caption{(a) Probability density of the predicted mean degree for
different values of species richness. The number of interactions was
predicted using the flexible links model fitted to all empirical
networks in our complete dataset. (b) Degree distributions of maximum
entropy for a network of \(27\) species and different numbers of
interactions. The numbers of interactions correspond to the lower and
upper bounds of the \(67%
\), \(89%
\), and \(97%
\) percentile intervals (PI), as well as the median, of the
counterfactuals of the flexible links model.}\label{fig:degree_dist_fl}
}
\end{figure}

@@

We define networks of maximum entropy as the configuration of the
adjacency matrix with the highest SVD entropy under a set of
constraints. As mentioned in Box 1, we used the SVD entropy as our
measure of entropy since it has been shown to be a reliable measure of
food-web complexity (Strydom, Dalla Riva, and Poisot 2021), in addition
to having the required properties of a proper measure of information
entropy. We thus seek to find the network with the highest complexity,
or randomness, that exactly reproduces specified constraints on its
structure. Our method is in contrast with maximum entropy graph models
that predict a probability distribution on networks under soft or hard
constraints (e.g., Park and Newman 2004; Cimini et al. 2019). We believe
our approach to be more flexible, easier to compute, while allowing
direct comparisons of empirical food webs with more complex networks
with similar structure.

@@

We compared different measures of the structure of empirical food webs
with the ones of null, neutral, and MaxEnt networks
tbl.~\ref{tbl:measures_all}, and \ref{tbl:measures_abund}. For instance,
we measured the nestedness \(\rho\) according to Staniczenko, Kopp, and
Allesina (2013), i.e.~using the spectral radius of the adjacency matrix.
Nestedness indicates how much the diet of specialist species is a subset
of the one of generalists, and is strongly associated with network
modularity (Fortuna et al. 2010). We also measured the maximum trophic
level \emph{maxtl}, network diameter \emph{diam} (i.e.~the longest of
the shortest paths between all species pairs; Albert and Barabasi 2002),
the average maximum similarity between species pairs \emph{MxSim}, the
proportion of cannibal species \emph{Cannib} (i.e.~the proportion of
self loops), and the proportion of omnivorous species \emph{Omniv}
(i.e.~species whose preys are of different trophic levels).
\emph{MxSim}, \emph{Cannib}, and \emph{Omniv} are more deeply defined in
Williams and Martinez (2000).

@@

Moreover, we found that empirical networks are close to their maximum
entropy considering a fixed joint degree sequence (fig.~S3). Empirical
networks had an SVD entropy of \(0.89 \pm 0.04\), compared to type II
MaxEnt networks which had values of \(0.94 \pm 0.03\). The relationship
between the SVD entropy of empirical food webs and their maximum entropy
configuration is plotted in the last panel of fig.~\ref{fig:measures}.
We also found no relationship between the difference in SVD entropy
between empirical and type II MaxEnt networks and species richness, the
number of interactions, or connectance (fig.~S4). Similarly, we found no
correlation between the difference in SVD entropy and the Jaccard
distance of empirical and type II MaxEnt networks.

@@

The principle of maximum entropy is a robust mathematical method of
finding least-biased probability distributions that have some desired
properties given by prior knowledge on a system. We first used MaxEnt to
predict the joint degree distribution and the degree distribution of
maximum entropy given known numbers of species and of interactions. We
found that the resulting joint degree distributions were more symmetric
than the ones of empirical food webs, which suggests that other
constraints might be needed to improve those predictions. We also used
MaxEnt to predict networks of maximum entropy with a specified
structure. These networks are the most complex, or random, given the
constraints used. Likewise, we found that knowledge of species richness
and of the number of interactions were not sufficient to reproduce many
aspects of network structure found in nature. However, a model based on
the entire joint degree sequence, i.e.~on the number of preys and
predators for each species, gave more convincing results. Our type II
MaxEnt network model yielded better or similar predictions than the type
II null model, also based on the joint degree sequence, for most
measures considered including the motifs profile. This suggests that the
joint degree sequence drives many aspects of network structure. Indeed,
considering our findings that empirical networks are close to maximum
entropy for a given joint degree sequence, our results suggest that
food-web topology is configured almost entirely randomly around these
marginal numbers of predators and preys.

Our method and results could be used for different purposes. First, they
could be used as first-order approximations of network structure when
only state variables are known. This could prove useful when predicting
network structure at large spatial scales, where few ecological
information is known at that scale. Second, they could be used as
informative priors in Bayesian analyses of the structure of ecological
networks (e.g., Cirtwill et al. 2019). Third, they could be used to make
better predictions of pairwise species interactions by constraining the
space of feasible networks, as discussed in Strydom et al. (2021).
Finally, they could be used as alternative null models of ecological
networks to better understand ecological mechanisms driving food-web
structure. In that case, our model might need to be slightly adapted to
give a probability distribution of Boolean networks (in contrast with
point estimates of maximum entropy networks).

@@

\leavevmode\hypertarget{ref-Cirtwill2019QuaFrac}{}%
Cirtwill, Alyssa R., Anna Eklof, Tomas Roslin, Kate Wootton, and
Dominique Gravel. 2019. {``A Quantitative Framework for Investigating
the Reliability of Empirical Network Construction.''} \emph{Methods in
Ecology and Evolution} 10 (6): 902--11.
\url{https://doi.org/10.1111/2041-210X.13180}.
